Introducing the \NAME Framework\\ for Efficient Intrusion Detection\\ in the Internet of Things

### 2014-10-15 paper.tex
evaluation updates to match new scope/wording

For the evaluation we constructed a small WSN consisting of three nodes: an
end-device, a router and a coordinator. The end-device and the router were
completely autonomous nodes, while the coordinator consisted of an XBee module,
directly connected to a computer via a USB break-out board, allowing serial
access via a terminal.

We started from a basic application that measures light intensity and reports
it to the coordinator. The application was written both manually and generated.
On top of this baseline we added two intrusion detection related algorithms: a
heartbeat, that allows nodes to validate each other's continuous presence, and
a reputation-building algorithm that checks if a parent node is cooperative and
actually forwards messages with a destination further down the
network\cite{ganeriwal2008reputation}.

@@

In case of the manual implementation, both algorithms were constructed as
standalone modules and sequentially called from the base-application's event
loop. The code generator was provided with FOO-lang descriptions of the
algorithms and generated all four cases.

Before reviewing the results, it is very important to put these results in a
proper perspective. Comparing manually written code to generated code is no
exact science. One can always argue that the quality of the two code bases is
not comparable and that the manual code always can be improved.

Still it might be interesting to look at the numbers, given that both code
bases are constructed with the same intentions and practices. We believe we
have achieved this and have taken honest decisions while constructing the
implementations, making them comparable. Numerical analysis of the
implementation offers us a preliminary estimation of the effect of our proposed
solution.

@@

In the case of the generated implementation, this no longer holds: although
that even the individual increases due to adding a single algorithm are higher
than in the manual case, the result for the implementation of both algorithms
is less than the sum of both. Here we clearly see the effect of reusing the
framework that comes with FOO-lang. The same goes for all other metrics. Maybe
most remarkable is the effect on the time of one event loop cycle: both
algorithms by itself add 150\% with respect to the base case, but when
combined, the impact hardly increases. Here we learn that the introduced
framework is responsible for the initial overhead, but it clearly pays for
itself when adding more algorithms.

@@

Secondly, we see that the framework that comes with the generated code adds to
the size of the resulting image. But the cost is almost constant and is
relatively lower when algorithms are combined. With roughly 3KB of additional
overhead, or 15\% in this simple case, this impact is affordable.

From a functional point of view, the generated code lives up to the
expectactions: both the number of frames as the bytes sent benefit from well
organized code and reuse of a common framework.

@@

Without any effort to optimize the framework code, nor leveraging knowledge
about the specific algorithms, this very basic generated code shows that the
proposed principles are not only theoretically feasible but actually result in
interesting improvements.

>>>>

For the evaluation we constructed a small meshed network consisting of three
devices, with only one device able to talk to both other devices, requiring
routing of messages through this device.

We started from a basic application that measures light intensity. The
application was written both manually and generated. On top of this baseline we
added two intrusion detection related algorithms: a heartbeat, that allows
nodes to validate each other's continuous presence, and a reputation-building
algorithm that checks if a parent node is cooperative and actually forwards
messages with a destination further down the
network\cite{ganeriwal2008reputation}.

@@

In case of the manual implementation, both algorithms were constructed as
standalone modules and sequentially called from the base-application's event
loop. The code generator was provided with \NAME descriptions of the algorithms
and generated all four cases.

Before reviewing the results, it is very important to put these results in a
proper perspective. Comparing manually written code to generated code is no
exact science. One can always argue that the quality of the two code bases is
not comparable and that the manual code always can be improved. Still it might
be interesting to look at the numbers, given that both code bases are
constructed with the same intentions and practices. We believe we have achieved
this and have taken honest decisions while constructing the implementations,
making them comparable. Numerical analysis of the implementation offers us a
preliminary estimation of the effect of our proposed solution.

@@

In the case of the generated implementation, this no longer holds: although
that even the individual increases due to adding a single algorithm are higher
than in the manual case, the result for the implementation of both algorithms
is less than the sum of both. Here we clearly see the effect of reusing the
software library that comes with \NAME. The same goes for all other metrics.
Maybe most remarkable is the effect on the time of one event loop cycle: both
algorithms by itself add 150\% with respect to the base case, but when
combined, the impact hardly increases. Here we learn that the introduced
software library is responsible for the initial overhead, but it clearly pays
for itself when adding more algorithms.

@@

Secondly, we see that the software library that comes with the generated code
adds to the size of the resulting image. But the cost is almost constant and is
relatively lower when algorithms are combined. With roughly 3KB of additional
overhead, or 15\% in this simple case, this impact is affordable.

From a functional point of view, the generated code lives up to the
expectations: both the number of frames as the bytes sent benefit from well
organised code and reuse of a common software library.

@@

Without any effort to optimise the software library, nor leveraging knowledge
about the specific algorithms, this very basic generated code shows that the
proposed principles are not only theoretically feasible but actually result in
interesting improvements.
